---
title: "Godfather Corleone"
author: "Michaela Fricova"
date: "2/9/2023"
output:
  html_document: default
  pdf_document: default
---

# The Corleone family network
This project is the first part of my mini-series on social network analysis. The goal of this series is to start with the very basics of network analysis, such as with the concepts of centrality and assortativity, and progress towards more advanced topics, including Erdős-Rényi and Configuration models, as well as Exponential family random graphs. In this first project, I am analyzing the network of **Corleone family**, as outlined in **The Godfather (1972)** film directed by Coppola and co-written by Mario Puzo. I depict the Corleone Sicilian mafia network based on characters' **on-camera dialogues/appearances**. For example, if Michael Corleone and Sonny Corleone have a dialogue in the movie, they are connected in this network. Since I am focusing on on-camera dialogues, the depicted network is undirected (i.e., there cannot be an arrow pointing from one character to the next as dialogues are bi-directional by their nature).

### Loading libraries
I start by loading the necessary libraries (igraph and igraphdata).

```{r error=FALSE, warning=FALSE, message=FALSE}
library(igraph)
library(igraphdata)
```

### Assigning edge attributes
I then create an **edgelist dataframe** of on-camera dialogues of the Sicilian mafia network (i.e., a dataframe of relationships in the network). For this, I am using the **graph()** function:

```{r}
mafia <- graph(edges=c(
  1,2,  1,3,  1,4,  1,5,  1,6,   1,7,  1,8,  1,9,  1,10,  1,11,  1,12,  1,13,  1,14,  1,15,  2,16,  2,17,  2,3,  2,18,  2,6,  2,19,  2,5,  2,20,  6,7,  6,21,  6,3,  6,22,  6,23,  6,24,  6,18,  6,10,  6,25,  6,26,  6,27,  6,28,  6,29,  6,20,  6,30,  6,9,  6,31,  6,32,  6,33,  3,21,  3,34,  3,10,  3,24,  3,4,  3,35,  8,11,  8,10,  10,11,  5,20,  5,9,  5,17,  9,34,  18,19,  19,32,  7,21,  7,30,  20,11,  20,12),  directed=FALSE)
```

### Assigning vertex attributes
After assigning these edge attributes of the network, I move on to assign **vertex/node attributes**. In other words, I assign names to each member of the network, their gender, their age and whether they survived till the end of the first Godfather movie. Additionally, I assign corresponding dead or alive status colors to each character for subsequent graphing purposes. I assign these node characteristics manually.

```{r}
# Character names
V(mafia)$name <-      c("Don Corleone","Sonny Corleone","Tom Hagen","Bonasera","C. C. Rizzi",
                       "Michael Corleone","Fredo Corleone","Luca Brasi","Johnny Fontane","Virgil Sollo",  
                       "Don Tattaglia","Don Barzini","Don Cuneo","Don Stracci","Anthony", 
                       "Lucy Mancini","Sonny's Wife","Peter Clemenza","Paulie Gatto","Carlo Rizzi",      
                       "Kay Adams","nurse","Enzo","Captain McCluskey","Don Tommasino",    
                       "Calo","Fabrizio","Don Vitelli","Apollonia","Moe Greene",       
                       "Willie Cicci","Rocco Lampone","Salvatore Tessio","Jack Woltz","Tessa")  

# Character status at the end of the film
V(mafia)$cue <-     c("deceased","deceased","alive","alive","alive",
                      "alive","alive","deceased","alive","deceased",
                      "deceased","deceased","deceased","deceased","alive",
                      "alive","alive","alive","deceased","deceased",
                      "alive","alive","alive","deceased","alive",
                      "alive","alive","alive","deceased","deceased",
                      "alive","alive","deceased","alive","alive")

# Vertex color associated with character status
V(mafia)$color <-   c("brown2","brown2","lightblue3","lightblue3","lightblue3",
                      "lightblue3","lightblue3","brown2","lightblue3","brown2",
                      "brown2","brown2","brown2","brown2","lightblue3",
                      "lightblue3","lightblue3","lightblue3","brown2","brown2",
                      "lightblue3","lightblue3","lightblue3","brown2","lightblue3",
                      "lightblue3","lightblue3","lightblue3","brown2","brown2",
                      "lightblue3","lightblue3","brown2","lightblue3","lightblue3")

# Character gender
V(mafia)$gender <-    c("M","M","M","M","F",
                        "M","M","M","M","M",
                        "M","M","M","M","M",
                        "F","F","M","M","M",
                        "F","F","M","M","M",
                        "M","M","M","F","M",
                        "M","M","M","M","M")

# Character estimated age in years
V(mafia)$age <-       c("60","35","35","60","25",
                        "25","30","40","40","60",
                        "60","60","60","60","2",
                        "25","30","40","50","40",
                        "30","40","30","45","60",
                        "30","30","55","20","50",
                        "35","40","45","45","45")

```

### Exploring network characteristics
I then proceed to explore the network. The following function **summary()** assures me that all 5 vertex attributes have been assigned correctly to the characters and that the network has been correctly labelled as undirected. 

```{r}
summary(mafia)
```

Then, the following functions tell me **the number of vertices** (35 in total) and **the number of realized edges** (60 in total), respectively:

```{r}
vcount(mafia)
ecount(mafia)
```

Finally, other descriptive functions are used also.

```{r}
degree(mafia)
mean(degree(mafia))
graph.density(mafia)
diameter(mafia)
```

They uncover that **Michael Corleone (21)** and **Don Corleone (14)** have the greatest **degrees**, i.e., the greatest number of on-screen dialogues. It seems women have dialogues much less frequently then men in the film. That can be further demonstrated through the **adjacency matrix** described in the following code section. 

```{r}
adj <- get.adjacency(mafia)
```

Overall, a character in the film is involved in approx. **3,43 degrees (dialogues)**. The network has **density** of 0.1008403 which implies that approximately 10.08% of all potential network dialogues are realized. All potential dialogues would be realised if each individual character would talk to each other individual character in the movie. The **diameter**, i.e. longest geodesic path length, is 3. That means that at most 3 movie characters are connected through their shared dialogues (i.e., one talks to a second and the second talks to a third character).

### Plotting the network
To plot the network graph, I am using the **ggraph()** function in this exercise. While doing so, I specify the size and color of each node, together with multiple other characteristics. The nodes are colour coded, according to the corresponding characters' survival status at the end of the movie - dark red/brown if they are deceased at the end of the movie, blue if they survived till the end. The vertex size corresponds to the degree of each character - the greater the number of on-screen dialogue a given character has, the bigger the size of their vertex bubble. The resulting network graph is below:

```{r}
E(mafia)$weight <- runif(ecount(mafia))
```

```{r}
library(ggraph)
library(graphlayouts)
degree_centrality <- degree(mafia, mode="all")
ggraph(mafia, layout = 'lgl') +
geom_edge_arc(color="gray", curvature=0.3) +
geom_node_point(color=V(mafia)$color, aes(size = degree_centrality)) +
geom_node_text(aes(label = name), size=3, color="gray40", repel=T) +
theme_void()+ theme(legend.position = 'none')
```








